---
title: "MAPS - Food Balance Sheet documentation"
author: "Lucia Segovia de la Revilla"
date: "Last compiled on `r format(Sys.time(), '%d %B, %Y')`"

output: html_document
---

# Food Balance Sheet

## Introduction


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

source(here::here("fbs.R"))

```

## Food composition data

1) Regional African FCT (Joy et al., 2014). Mainly minerals
2) Kenya FCT (FAO/Government of Kenya, 2018). It has matches for all food dictionary codes by `r count(distinct(fbs %>% select(1,4:5) %>% distinct() %>% 
  left_join(., MAPS_ken) %>% filter(is.na(original_food_id))))`
3) UK FCT

## Data issues

**Wrong coding**

1) Marine fish, other was wrongly coded to 2763, which is pelagic fish.

2764 = Marine Fish, Other
2763 = Pelagic Fish

**Items hardly consumed**

1) 2562 - palm kernels (1491.02.01): the mean value of consumption is zero `r mean(fbs$amount_consumed_in_g[fbs$food_genus_id=="1491.02.01"])`. It is hardly reported in any FCT, as it is not consumed but used to produce oil and other products. 


```{r cars}
summary(cars)
```




  

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
