---
title: "Ethiopia HCES food matching"
author: "Segovia D L Revilla, Lucia"
date: "03/03/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Matching to dictionary

Check Notes

1. Teff needed to be added to dictionary.
2. Wheat - what should be matched to? - refined and unrefined grain?
3. Barley - new to dictionary & check how is consumed (unrefined?)
4. Maize - how is consumed (yellow and white grain)
7. Horsebean = broad bean - new to dictionary. Is it dried or fresh?
8. Field peas = dried peas
9. chickpeas - new to dictionary
9. Other pulse or nut (SPEC --> wrong coding. We should recode it to 9 = chickpeas.
10. Lentils - what type?
11. Haricot bean - Double-check variety
12. Niger seeds - New to dictionary
17. kocho - bread-like food from ensete - new to dictionary. We put it under roots and tuber (ref. Food balance sheet handbook, 2001, page 70) 
20. Cheese - we used cottage cheese because Ethiopian cheese is similar.
25. Chat / Kat - Not sure if this is food..
24. Bula - ensete flour - new to dictionary 
60. Other cereal (SPECIFY) - should check
111. Other pulse or nut (SPEC - should check
131. Other seed (SPECIFY) - check
143. "Greens" - cabbage and spinach, but should check what are the mean greens!
145. Other vegetable (SPECIFY - check
152. Other fruit (SPECIFY) - check
170. Sweet potato - white flesh
173. Godere = taro
174. Other tuber or stem (SPE - check (low no.)
183. Fish - very low consumption (only reported 99 times) - used tilapia (ref. FAO 2022. Fishery and Aquaculture Country Profiles. Ethiopia. Country Profile Fact Sheets. Fisheries and Aquaculture Division [online]. Rome. [Cited Thursday, March 3rd 2022]. https://www.fao.org/fishery/en/facp/eth?lang=en )
195. PuUrchased Injera - new to dictionary
196. "Purchased Bread or Biscu" - bread, refined flour and biscuits - check
198. Other prepared food and - check
202. Oils (processed) - palm oil. Acc. to publications (Sheehy, et al., 2019; FAO, 2008) and FBS.
206. Tella - traditional beer made from maize or wheat. (https://doi.org/10.1016/j.jef.2015.08.002)


```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
