---
title: "Ethiopia HCES food matching"
author: "Segovia D L Revilla, Lucia"
date: "03/03/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Matching to dictionary

### ESS3

Check Notes

1. Teff needed to be added to dictionary.
2. Wheat - what should be matched to? - refined and unrefined grain?
3. Barley - new to dictionary & check how is consumed (unrefined?)
4. Maize - how is consumed (yellow and white grain)
7. Horsebean = broad bean - new to dictionary. Is it dried or fresh?
8. Field peas = dried peas
9. chickpeas - new to dictionary
9. Other pulse or nut (SPEC --> wrong coding. We should recode it to 9 = chickpeas.
10. Lentils - what type?
11. Haricot bean - Double-check variety 
12. Niger seeds - New to dictionary
17. kocho - bread-like food from ensete - new to dictionary. We put it under roots and tuber (ref. Food balance sheet handbook, 2001, page 70) 
20. Cheese - we used cottage cheese because Ethiopian cheese is similar.
25. Chat / Kat - Not sure if this is food..
24. Coffee - beverage, but check
26. Bula - ensete flour - new to dictionary 
60. Other cereal (SPECIFY) - should check
111. Other pulse or nut (SPEC - should check
131. Other seed (SPECIFY) - check
143. "Greens" - cabbage and spinach, but should check what are the mean greens!
145. Other vegetable (SPECIFY - check
152. Other fruit (SPECIFY) - check
170. Sweet potato - white flesh
173. Godere = taro
174. Other tuber or stem (SPE - check (low no.)
181. Beef, used w/o bones. 
183. Fish - very low consumption (only reported 99 times) - used tilapia (ref. FAO 2022. Fishery and Aquaculture Country Profiles. Ethiopia. Country Profile Fact Sheets. Fisheries and Aquaculture Division [online]. Rome. [Cited Thursday, March 3rd 2022]. https://www.fao.org/fishery/en/facp/eth?lang=en )
195. PuUrchased Injera - new to dictionary
196. "Purchased Bread or Biscu" - bread, refined flour and biscuits - check
198. Other prepared food and - check
202. Oils (processed) - palm oil. Acc. to publications (Sheehy, et al., 2019; FAO, 2008) and FBS.
203. Tea - matched to beverage - check
206. Tella - traditional beer made from maize or wheat. (https://doi.org/10.1016/j.jef.2015.08.002)

### ESS4

207. Vetch = *Vicia Sativa* (acc. to FBS handbook) (Fava bean/ broad bean) - new to dictionary.
208. Fenugreek - new to dictionary
209. mung bean = *Vigna ssp.* (acc. to FBS handbook) new to dictionary
210. Processed pulses (Shiro) - chick peas flour - doble-check
404. kale - add to dictionary
404. cabbage - cabbages, white, raw
406. Garlic - add to dictionary
407. Moringa - add to dictionary
608. Carrot - add to dictionary
609. Beetroot - add to dictionary
807. Gesho (hops stem) - add to dictionary


```{r cars}
summary(cars)
```

## Matching with KENFCT

There are 29 food items w/o a match. We are checking whether they are available or nor in KENFCT.

List of items not found and possible exchange (if any)

1. There is no refined wheat grain
2. There is no barley
3. Haricot beans, lima bean
4. Niger seeds
5. Linseed 
6. Kocho 
7. Coffee, beverage, instant coffee powder (Check what type of coffee and units)
8. Bula
9. Sheep/mutton meat
10. Injera
11. Oil - only red palm oil
12. Soft drinks/Soda
13. Beer
14. Tella

Need to change code
 1. Chicken - was wrong code. Need to fix it from fuzzy_matching input



```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
