---
title: "fct_cleaning"
author: "LuciaSegovia"
date: "8/12/2020"
output:
  pdf_document: 
    latex_engine: xelatex
  html_document:
    df_print: paged
  word_document: default
---

```{r setup, include=FALSE, warning=FALSE}

knitr::opts_chunk$set(echo = TRUE,warning=FALSE, message=FALSE) 

knitr::read_chunk('fct_cleaning.R')

library(tidyverse)

```

First, we are loading the tidyverse package and the FCT_QA spreadsheet. This spreadsheet is where all the information about each FCT is gathered. 

Then, we are going to explore all the FCT in excel. We need to identify if the file has more than one worksheets and, if it does, where FCT is located. 

...NOTE:This is something where I need to improve this bit of code

```{r, WAFCT-tagname, echo=FALSE}
```

```{r exploring sheets - loop}

#Loading FCT_QA data set

FCT_QA <- readxl::read_excel(here::here('data', 'FCT_QA.xlsx'), sheet = 2)

##TRIAL 1: This loop only works if all statements are = TRUE

sheet <- vector('double',  nrow(FCT_QA))

for(e in 1:nrow(FCT_QA)) {
  x <- paste(paste(FCT_QA[e,6]
                   , FCT_QA[e,2], sep = '_'), 'xlsx', sep = '.')
  
  sheet[[e]] <- readxl::excel_sheets(here::here( 'data',
                                                x)) %>% list()
}

sheet

```

Once we know what it is in each spreadsheet, we can load those sheets of interest. For instance, for West Africa FCT we will use the worksheet 5 ('05_sum_57 (per 100g EP)'). While for Kenya FCT, we need worksheet 4.

Since we want to perform some analysis, we need to standardize the variables. This process is also valuable to analyze the quality of each FCT and to record potential issues or aspects that need to be kept in mind when working with different FCT. Such as, are all same variables calculated in the same way? Are they comparable/ interconverteable? Are same variables reported in the same units? Are our variables of interest reported? 

Also, to run quality check such as, recalculation of variables, SOP, etc.

For this section, we are going to calculate the missing values per MN of interest and the Sum of Proximates (SOP) calculated as the sum of water, protein, fat, available carbohydrates, dietary fibre (or total carbohydrates), alcohol (if applicable) and ash. 

SOP = SUM of WATER + PROTCNT + FAT + CHOAVLDF + FIBFG + ALC + ASH. 

In order to do so, we are going to standardize variables names as descibed above for food items names and code and for food components tagnames (FAO/INFOODS). Additionally, food groups will be standardized using FoodEx2 classification system.

####Standtadization of variables using FAO/INFOODS tagname

####1) West Africa FCT

Since WAFCT is very recent and it is been done according to FAO/INFOODS tagname, we do not need to standardize much of the variables. Only for code, fooditem, fooditemFR, scientificName, and ref.

UPDATE: Add rename to Energy variables following the standard that it is set below (MAFOODS) 
 +ENERC1 = ENERC...10 = Energy in kcal; 
 +ENERC2 = ENERC...10 = Energy in kJ).
 
When different way of reporting same variable (i.e. dietary fibre vs crude fibre), most of the time they were reported between brackets. We have desegregated those values to be different variables (columns). Then, we can decide which values we want to use (hierarchical decision). It is also good pratice to visualize the real number of entries for a particular component. Particularly relevant for

FOL: some of the values are reported as FOLSUM. This tagname is recommended to be discontinued. 

PHYYCPP: some values were reported as PHYTCPPD or PHYCPPI.

There is an issue with values that are between [], when converting the variable into numeric those values are coerced as NA. There are two options here:
 
 1) to eliminate the [] and to use the values: Then, we are loosing the reason of putting them between [] which is to remark the low quality of that entry.
 2) to allow them to be converted in NA: Hence, we are not using low quality information.
 
We do not want to loose information but also we want to use the highest quality available. 

A piece of code is written to remove food groups from the code column and convert it in a new variable. This will be useful for subsetting by 'foodgroups'. I.e. cereals and other staples. That part will be reconverted when food groups are standardized according to FoodEx2.

UPDATE: We have updated SOP calculation to account for FAT and FATCE.


 
```{r WAFCT standard variables}

##View WAFCT structure

readxl::read_excel(here::here('data', 
                              '2019_WAFCT.xlsx'), sheet = 5) %>% head()


##Loading WAFCT, skip 2 first rows to use tagnames for components

WAFCT <- readxl::read_excel(here::here( 'data', 
                                       '2019_WAFCT.xlsx'), sheet = 5, skip = 2) %>%
  mutate(FCT = 'WAFCT') %>% glimpse()


#rename variables according to our standards

WAFCT <- WAFCT %>% rename(code = '...1', 
                          fooditem = '...2', 
                          fooditemFR = '...3',
                          scientificName = '...4',
                          ref = '...5' ,
                          ENERC2 = 'ENERC...9', 
                          ENERC1 = 'ENERC...10') 


#creating variable 'foodgroups'

#Extracting variables names only in English

fgwa <- WAFCT %>% filter(is.na(fooditem), !is.na(code)) %>% pull(code) %>%
  stringr::str_split_fixed( '/', n = 2) %>% as_tibble() %>% pull(V1)


WAFCT <- WAFCT %>% 
  mutate(foodgroup = ifelse(grepl("01_", code), fgwa[1],
                            ifelse(grepl("02_", code), fgwa[2],
                                   ifelse(grepl("03_", code), fgwa[3], 
                                          ifelse(grepl("04_", code), fgwa[4], 
                                                 ifelse(grepl("05_", code), fgwa[5],
                                                        ifelse(grepl("06_", code), fgwa[6],
                                                               ifelse(grepl("07_", code),  fgwa[7],
                                                                      ifelse(grepl("08_", code), fgwa[8],
                                                                             ifelse(grepl("09_", code), fgwa[9], 
                                                                                    ifelse(grepl("10_", code), fgwa[10], 
                                                                                           ifelse(grepl("11_", code), fgwa[11],
                                                                                                  ifelse(grepl("12_", code),  fgwa[12],     
                                                                                                         ifelse(grepl("13_", code), fgwa[13],
                                                                                                                ifelse(grepl("14_", code), fgwa[14],
                                                                                                                       'NA'))))))))))))))) %>% 
  filter(!is.na(fooditem))

#Extracting variables calculated with different method and reported as using []


#This keeps [] 

#WAFCT <- WAFCT %>% mutate(FIBC = str_extract(FIBTG, "\\[.*?\\]"))


#This keeps only numbers

WAFCT <- WAFCT %>% mutate(FIBC = str_extract(FIBTG, '(?<=\\[).*?(?=\\])')) %>%
  mutate(FATCE = str_extract(FAT, '(?<=\\[).*?(?=\\])')) %>%
  mutate(TOCPHA = str_extract(VITE, '(?<=\\[).*?(?=\\])')) %>%
  mutate(FOLSUM = str_extract(FOL, '(?<=\\[).*?(?=\\])')) %>%
  mutate(PHYTCPPD_I = str_extract(PHYTCPP, '(?<=\\[).*?(?=\\])'))

#Reordering variables

WAFCT <- WAFCT %>% dplyr::relocate(foodgroup, .after = ref) %>%
  dplyr::relocate(FCT, .before = code)

#WE don't need this bit anymore
#WAFCT <- WAFCT[,c(64, 1:4, 65, 5:69)]


#Converting into numeric numeric variables  

WAFCT<- WAFCT %>% mutate_at(vars(8:69), funs(as.numeric)) 

#Calculating SOP for the WAFCT


WAFCT <- WAFCT %>%
  mutate(SOP = ifelse(is.na(FATCE),
    reduce(select(.,
                             'WATER', 'PROTCNT' ,'FAT', 'CHOAVLDF','FIBTG', 'ALC', 'ASH'), `+`),
    reduce(select(.,
                  'WATER', 'PROTCNT' ,'FATCE', 'CHOAVLDF','FIBTG', 'ALC', 'ASH'), `+`)))

summary(WAFCT$SOP)


```

NOTE: The problem with converting to numeric is that number between [] are removed (as NA). It is also creating a issue with the calculation of SOP

####2) Malawi FCT

For Malawi dataset all the variables names need to be converted manually because in this FCT component names are not standardized using tagnames or other method. To rename all the component in a standardized way three parameters needs to be taken into account:

          1) component name
          2) unit reported
          3) method used for analysis or calculation. 
      I.e. for total nitrogen reported in grams and calculated with Kjeldahl method the tagname is NT

The guidelines used to select the tagnames are 'FAO/INFOODS Guidelines for Checking Food Composition Data prior to the Publication of a User Table/Database Version 1.0, 2012', 'FAO/INFOODS Guidelines Guidelines for Converting Units, Denominators and Expressions Version 1.0, 2012' and the tagnames found in FAO/INFOODS website starting from the most recent (2010 tagnames > 2008 tagnames > 2007 tagnames). 

Decisions for tagnames in Malawi dataset

1) Energy reported as kcal (ENERC1), energy reported in kJ (ENERC2)

2) Nitrogen (g) = NT because it is reported in grams and it is mainly calculared by Kjeldahl method

3) Total protein (g) =  PROTCNT. Most of the values were calculated from total, nitrogen, In addition, following the advice on FAO/ INFOODS. 'PROCNT' is the appropriate tagname for total protein in most cases.

4) Total fat (g) = FAT (fat, total, Unit: g, Synonyms: total lipid). There are a mix of method on how fat has been calculated, including Soxhlet method that tends to underestimate content. However, since we are not currently differentiating between methods we will use the most generic version of the tagname. 
NOTE: in the future we could use FAT for normal and FATCE for continuous extraction 

5) Saturated FA (g) = FASAT (fatty acids, total saturated, Unit: g)

6) Mono-unsaturated FA (g) = FAMS (fatty acids, total monounsaturated, Unit: g)

7) Polyunsaturated FA (g) = FAPU (fatty acids, total polyunsaturated, Unit: g)

8) Cholesterol (mg) = CHOLC (cholesterol; determined by chemical method (classical), Unit: mg). We don't have information on how this variable has been obtained and we have chosen the 'classial' version. 
NOTE: This tagname could be updated in the future.

9) Total CHO for UDB = CHOCSM (carbohydrate, total; calculated by summation, Unit: g, Comments: This value is the sum of the, sugars, starches, oligosaccharides, and carbohydrate dietary fibre.). It was the only total CHO by summation. (CHOALVDF + Total Fibre)

10) Carbohydrate, avail. (g) = CHOAVLDF (This value is calculated: 100 g minus total grams of water, protein, fat, alcohol, ash and dietary fibre; or CHOCDF – dietary fibre.)

11) Total sugar (g) = SUGAR (sugars, total, Unit: g, Comments: Sum of free monosaccharides and disaccharides.) From 2010 review tagnames, it is also suggested for sugars when the the method is unknown. 

12) Added sugar (g) = SUGAD (added sugar, Unit: g)

13) Total fibre (g) = FIBC (fibre, crude, Unit: g Comments: The crude fibre method of fibre analysis is obsolete.). Other methods were used as well. 
NOTE: This term needs to be reviewed

14) Starch (g) = STARCH (starch, total, Unit: g, Comments: The sum of all polysaccharides yielding glucose after hydrolysis with suitable enzymes; includes amylose, amylopectin, glycogen, and dextrins.The expression is unknown if by weight or in monosaccharide equivalent (2010)).

15) Ash (g) =  ASH (ash, Unit: g, Synonyms: minerals)

16) Ca (mg) = CA (calcium, Unit: mg)

17) Fe (mg) = FE (iron, total, Unit: mg, Comments: Includes both haem and non-haem iron.). We assume total iron since we have no extra information.

18) Mg (mg) = MG (magnesium, Unit: mg). This was the only entry available.

19) P (mg) = P (phosphorus, Unit: mg). This was the only entry available.

20) K (mg) = K (potassium, Unit: mg). This was the only entry available.

21) Na (mg) = NA (sodium, Unit: mg). This was the only entry available. 

22) Zn (mg) = ZN (zinc, Unit: mg). This was the only entry available.

23) Cu (mg) = CU (copper, Unit: mg). This was the only entry available.

24) Mn (mcg) = MN (manganese, Unit: mcg). This is the only entry available.

25) I (mcg) = ID (iodine, Unit: mcg). This is the only entry available.

26) Se (mcg) = SE (selenium, Unit mcg). This is the only entry available.

27) Vitamin A (RAE) (mcg) = VITA_RAE (Total vitamin A activity expressed in mcg retinol activity equivalent, Unit: mcg,  Comment RAE = mcg retinol+ 1/12 mcg ß- carotene + 1/24 mcg other provitamin A carotenoids (or RAE= mcg retinol + 1/12 mcg ß- carotene equivalent)

28) Vitamin A (RE) (mcg) = VITA (vitamin A; calculated by summation of the vitamin A activities of retinol and the active carotenoids, Unit: mcg. Synonyms: retinol equivalents. Comments: Total vitamin A activity = mcg retinol + 1/6 mcg beta-carotene + 1/12 mcg other provitamin A carotenoids.) As per the calculated in MAFOODS ((Retinol, µg/100g EP × 1) + (β-carotene, µg/100g EP ÷ 6) + (α-carotene, µg/ 100g EP ÷ 12) + (β-cryptoxanthin, µg/100g EP ÷ 12)).

29) Thiamin (mg) = THIA (thiamin, Unit: mg, Synonyms: vitamin B-1; aneurin; thiamine)

30) Riboflavin (mg) = RIBF ( riboflavin, Unit: mg, Synonyms: Vitamin B-2; riboflavine.). This is the only entry available.

31) Niacin (mg) = NIA (niacin, preformed, Unit: mg, Synonyms: nicotinic acid; nicotinamide (These terms are not true synonyms but are often used in food tables to refer to niacin.)). Since, we have no more information, we assume that it is niacin preformed. We could calculate NIAEQ if we have information tryptophan.

32) Vitamin B6 (mg) = VITB6 (vitamin B-6, total; method of determination unknown variable, Unit: mg, Comments: The <VITB6-> tagname should be used if it is not known whether  the vitamin B-6 value was determined by analysis or by calculation.). We use the generic version duet to lack of information on the method on how it has been calculated. 

33) Folic acid (mcg) = FOL (folate, total, Unit: mcg, Synonyms: folacin; folic acid, Comments: Includes both conjugated and free folate, Determined by microbiological assay)). This is the recommended expression. However, more information on how this component has been calculate should be provided.  
Note: Review this content with recently updated info. 

34) Vitamin B12 (mcg) = VITB12 (vitamin B-12, Unit: mcg, Synonyms: cobalamin, Comments: Includes all the active forms of vitamin B-12 in food.). This is the only entry available.

35) Pantothenate (mg) = PANTAC (pantothenic acid, Unit: mg, Synonyms: D-pantothenate; vitamin B-5 (obsolete term, may still appear in some tables)). This is the only entry available.

36) Biotin (mcg) = BIOT (biotin, Unit: mcg, Synonyms: vitamin H). This is the only entry available.

37) Vitamin C (mg) = VIT C (vitamin C, Unit: mg, Synonyms: ascorbic acid; ascorbate  (Note that these terms are not true synonyms but are often found in food tables to refer to vitamin C.), Comments: L-ascorbic acid plus L-dehydroascorbic acid.). We used this tagname because most of the method used to calculate vit. C were to calculate both fractions (total vit. C), as HPLC and spectro-photometric method. For the two other items that DCPIP was used, this were raw version of the food items, for this case vit C should be similar. Although it might be worth to check the impact of milling on L-ascorbic acid.

38) Vitamin D (mcg) = VITD (vitamin D; calculated by summation of ergocalciferol and cholecalciferol, Unit: mcg. Synonyms: calciferol, Comments: Ergocalciferol plus cholecalciferol.). Here, we should be using VITD- (vitamin D; method of determination unknown or variable, Comments: The <VITD-> tagname should be used if it is not known whether the vitamin D value was determined by bioassay or by calculation.)

39) Vitamin E (mg) = VITE (vitamin E; calculated by summation of the vitamin E activities of the active tocopherols and tocotrienols; expressed as alpha-tocopherol equivalents Unit: mg.). As with VITD, VITE- should be used as the method is unknown. 

40) Phytic Acid (mg) = PHYT (Phytic acid, unknown or variable method, Unit: mg). This should be PHYT-. It is the most generic version of phytic acid, since we have very little information on how it has been calculated.
Note: information on tagnames for phytates are in the S. Dahdouh et al. 2019. 


Note: Some of the generic tagname should be followed by '-' (dash), however since this character could cause some issues with the code, we have dropped the '-'.


```{r MAFOODS tagname}

##View FCT structure
#Creating a variable for the names of the data-set 
# it is composed by 2 columns on the FCT_QA dataset
#Column 6 == Year
#Column 2 == Short_name
#Just need to input the number of the row (1-13)

x <- 2


readxl::read_excel(here::here( 'data', 
                              paste(paste(FCT_QA[x,6],
                                          FCT_QA[x,2], sep = '_'), 'xlsx', sep = '.')), sheet = 1) %>% head()


MAFOODS <- readxl::read_excel(here::here( 'data', 
                                         paste(paste(FCT_QA[x,6],
                                                     FCT_QA[x,2], sep = '_'), 'xlsx', sep = '.')), 
                              sheet = 1) %>% mutate(FCT = 'MAFOODS') %>% glimpse()

#Renaming variables with tagnames and converting numeric variables into numeric

FCT_tag <- c('code', 'ref', 'fooditem', 'foodgroup', 'WATER', 'ENERC1', 'ENERC2', 'NT',
             'PROTCNT', 'FAT', 'FASAT', 'FAMS', 'FAPU', 'CHOLC', 'CHOCSM', 'CHOAVLDF',
             'SUGAR', 'SUGAD', 'FIBC', 'STARCH', 'ASH', 'CA', 'FE', 'MG', 'P', 'K', 'NA',
             'ZN', 'CU', 'MN', 'ID', 'SE', 'VITA_RAE', 'VITA', 'THIA', 'RIBF', 'NIA', 'VITB6', 'FOL', 
             'VITB12', 'PANTAC', 'BIOT', 'VITC', 'VITD', 'VITE' ,'PHYT', 'FCT')


MAFOODS <- MAFOODS %>% rename_all( ~ FCT_tag) %>% mutate_at(vars(5:46), funs(as.numeric)) 

#Calculating SOP - Removing ALC and changing FIBTG for FIBC

MAFOODS <- MAFOODS %>%
  mutate(SOP = reduce(select(.,
                             'WATER', 'PROTCNT' ,'FAT', 'CHOAVLDF','FIBC',  'ASH'), `+`))

summary(MAFOODS$SOP)

```

####3) Ethiopia FCT

5) Moisture (%): WATER (Unit: g). It is reported in %. 

9) CHO (inc. fiber) (g): CHOT (total carbohydrates, Unit: g). It is reported including fibre, hence it is 'total CHO'. Because the method is unknown we use the generic CHOT. 

10) Fiber (g): FIBTG (dietary fibre, Unit: g). We are going to assume that it is according to the standards and it is dietary fibre, until further clarification can be obtained. 

16) Beta-carotene Equiv. (mcg): CARTBEQ (beta-caroten equivalent, Unit: mcg). It is not recommended to be listed alone in the FCT. Since we are going to use VITA_RAE or VITA when possible. We can convert this variable to VITA_RAE and VITA by using the following equations. However, retinol information is missing in the ETHFCT.  

Eq.4:  VITA_RAE = mcg retinol + 1/12 mcg ß- carotene equivalent

Eq.5:  VITA = mcg retinol + 1/6 mcg ß- carotene equivalent

20) Tryptophan (mg) : TRP (L-tryptophan, Unit: mg)

21) Ascorbic acid (mg): We are going to assume that it is VitC. 

22) Refuse (%): REFUSE. It might be re-converted to EDIBLE portion to keep consistency

NOTE:Still need to arrange foodgroups

NOTE: When calculating SOP, we do not include fibre because it is already included in CHOT.

####4) Gambia FCT

Within the GMBFCT there are tagnames assigned to each food component

4) Energy...4(kcal): ENERC1 (it should be ENERGYC, as per GMBFCT), this should be re-calculated as CHOAVL but it is not possible since we have no information to convert monosacharides information to weight. Hence, we are going to use it as it is and compared it with the other ENERC.

5) Energy...5 (kJ): ENERC2 (it should be ENERGYC, as per GMBFCT), this needs to be re-calculated as per kJ contribution per macronutrient.

8) Carbo-hydrate (g): CHOAVLM (Carbohydrates availables expresed as monosacharides equivalents, Unit: g). This can not be converted due to lack of information of the carbohydrates composition and can not be compared with available CHO in weight.

9) Fibre (g): FIBTS (total dietary fibre, sum of non-digestible polysaccharides and lignin, Unit: g). More information is needed to know how the estimates are affected by the method. 

10) Phytate (g): PHYTAC. I think it should be PHYPPCD - Phytic acid, based on phytate phosphorus estimated by direct ferric precipitation (Unit: mg) because it used FE precipitation. 

16) Carotene (mcg): CARBEQ (beta-caroten equivalent, Unit: mcg). As in ETHFCT, this should not be listed alone in the FCT.

Ash are not reported, we can see a slightly lower SOP when comparing it with MAFOODS and WAFCT.

```{r GBMFCT tagname }

#Gambia FCT - tagname standardization

#View FCT structure
#Creating a variable for the names of the data-set 
# it is composed by 2 columns on the FCT_QA dataset
#Column 6 == Year
#Column 2 == Short_name
#Just need to input the number of the row (1-13)

x <- 4


readxl::read_excel(here::here( 'data', 
                              paste(paste(FCT_QA[x,6],
                                          FCT_QA[x,2], sep = '_'), 'xlsx', sep = '.')), sheet = 1) %>%
  head()

#Customized saving FCT

GMBFCT <- readxl::read_excel(here::here( 'data', 
                                        paste(paste(FCT_QA[x,6],
                                                    FCT_QA[x,2], sep = '_'), 'xlsx', sep = '.')), 
                             sheet = 1) %>% mutate(FCT = 'GMBFCT') %>% glimpse()


FCT4_tag <- c('foodgroup','code',  'fooditem', 'ENERC1', 'ENERC2',  
              'PROTCNT', 'FAT',  'CHOAVLM', 'FIBTS', 'PHYTAC',
              'WATER', 'CA',  'P', 'FE',  'ZN', 'CARTBEQ', 'VITC',
              'ref', 'FCT')

#Renaming and converting into numeric (needed for binding datasets)

GMBFCT <- GMBFCT %>% rename_all( ~ FCT4_tag) %>% slice(3:527) %>%
  mutate_at(vars(4:17), funs(as.numeric)) 

#Converting into character (needed for binding datasets)

GMBFCT$code <- as.character(GMBFCT$code)

#SOP - changing CHOAVLDF to CHOAVLM, FIBTG to FIBTS, removing ALC, no ASH are included.

GMBFCT <- GMBFCT %>%
  mutate(SOP = reduce(select(.,
                             'WATER', 'PROTCNT' ,'FAT', 'CHOAVLM','FIBTS'), `+`))


```


####5) Kenya FCT

Within this FCT there are a set of components with tagnames (majority of items not relevants for the project and with high amount of missing values) and other without. To facilitate the standardization of relevant component, we have cut the table in two

KENFCT1 - We have review and update tagnames and we have cut the rows until 1242 to be consistent with the number of food items reported. The two extra-rows removed contains information on FOLAC for some food items and an the following statement '**values between [ ] were borrowed from an average of blood from other animals'. Then tagnames were standardized following the standard names as described above and on FCT_QA, some decision point are described below: 

22) Retinol (mcg): RETOL (Retinol, Unit: mcg)

27-28) Folates group: According to the User-guide all four forms were reported (Dietary folate eq., food folate, synthetic folate, total folate). However when exploring the variables only two are present (dietary folate eq., and food folate). FOLAC is reported as comments or food notes for those food items that where fortified. An control check should be implemented for fortified food, since some of the fortified food (FOLDFE ≠ FOLDF) are not always stated in the 'fooditem' (name). FOL was not found in the user-guider nor in the xlsx format. 

KENFCT2 - We have incorporated tagnames as per in the dataset (with no modifications)

####6) Lesotho FCT

4) Energy is only reported in kJ. New variable need to be created, for ENERC1.

We need to clean the table from missing values that were originated by the format of the table (double-spaces)

NOTE: I need to update the loop for foodgroup variable. None of the loops that I tried worked so far...

```{r LSOFCT tagname }

#View FCT structure
#Creating a variable for the names of the data-set 
# it is composed by 2 columns on the FCT_QA dataset
#Column 6 == Year
#Column 2 == Short_name
#Just need to input the number of the row (1-13)

x <- 6


readxl::read_excel(here::here(  'data', 
                                paste(paste(FCT_QA[x,6],
                                            FCT_QA[x,2], sep = '_'), 'xlsx', sep = '.')), sheet = 6) %>%
  head()

#Customized saving FCT

LSOFCT <- readxl::read_excel(here::here(  'data', 
                                           paste(paste(FCT_QA[x,6],
                                                       FCT_QA[x,2], sep = '_'), 'xlsx', sep = '.')), 
                              sheet = 6) %>% mutate(FCT = 'LSOFCT') %>%  select(2:41) %>% glimpse()

FCT6_tag <- c('code', 'fooditem', 'EDIBLE', 'ENERC2', 'WATER', 
               'CHOAVLDF', 'NT', 'PROTCNT', 'PROPLA', 'PROANI' , 'FIBTG', 'ASH', 'FAT',
              'CHOLE', 'FASAT', 'FAMS', 'FAPU', 'STARCH', 'SUGAR',
              'CA', 'FE', 'MG', 'P', 'K', 'NA', 'ZN','CU','SE', 'MN', 
              'VITA',  'CARBEQ', 'VITD', 'VITE', 
              'THIA', 'RIBF', 'NIA', 'VITB6'  ,'FOL', 'VITC', 'FCT')

LSOFCT <- LSOFCT %>% rename_all( ~ FCT6_tag) 

#creating variable 'foodgroups'

#Extracting variables names only in English

fglso <- LSOFCT %>% filter(is.na(code), !is.na(fooditem)) %>% pull(fooditem) %>%
  stringr::str_split_fixed( ' ', n = 2) %>% as_tibble()

nolso <- fglso %>%  pull(V1)

fglso <- fglso %>%  pull(V2)


LSOFCT <- LSOFCT %>% 
  mutate(foodgroup = ifelse(grepl(nolso[1], code), fglso[1],
                            ifelse(grepl(nolso[2], code), fglso[2],
                                   ifelse(grepl(nolso[3], code), fglso[3],
                                    ifelse(grepl(nolso[4], code), fglso[4], 
                                  ifelse(grepl(nolso[5], code), fglso[5], 
                                ifelse(grepl(nolso[6], code), fglso[6], 
                                 ifelse(grepl(nolso[7], code), fglso[7],
                                ifelse(grepl(nolso[8], code), fglso[8], 
                                ifelse(grepl(nolso[9], code), fglso[9],
                                  ifelse(grepl(nolso[10], code), fglso[10], 
                                ifelse(grepl(nolso[11], code), fglso[11],
                                ifelse(grepl(nolso[12], code), fglso[12], 
                                ifelse(grepl(nolso[13], code), fglso[13], 
                                ifelse(grepl(nolso[14], code), fglso[14],
                                ifelse(grepl(nolso[15], code), fglso[15],
                              ifelse(grepl(nolso[16], code), fglso[16],
                                  ifelse(grepl(nolso[17], code), fglso[17],  
                                       ifelse(grepl(nolso[18], code), fglso[18],
                                      ifelse(grepl(nolso[19], code), fglso[19],
                                          ifelse(grepl(nolso[20], code), fglso[20], 
                                                           'NA')))))))))))))))))))))%>% 
                                               filter(!is.na(code))

LSOFCT <- LSOFCT %>%   mutate_at(vars(3:39), funs(as.numeric)) 


LSOFCT <- LSOFCT %>%
  mutate(SOP = reduce(select(.,
                             'WATER', 'PROTCNT' ,'FAT', 'CHOAVLDF','FIBTG',  'ASH'), `+`))

summary(LSOFCT$SOP)

```


####7) Mozambique FCT

ONLY PDF


####8) Nigeria FCT

All the variables are already labelled following FAO/INFOODS, hence only minor cleaning is needed. For MN, we need to capitalize them

13) FATCE: Fat values derived from Soxhlet method (Unit: g)

14) CHOCDF: total carbohydrate calculated by difference (Unit: g). This value is calculated: 100 - (weight in grams [water + protein + fat + ash + alcohol] in 100 g of food). In line with previous tagnames, we have changed 

15) FIB: We are going to assume that it is FIBGT, which is the recommended choice by FAO/INFOODS until we have more information. 

There are an issue with some values of protein. 


```{r NGAFCT tagname }

#Nigeria FCT - tagname standardization

#View FCT structure
#Creating a variable for the names of the data-set 
# it is composed by 2 columns on the FCT_QA dataset
#Column 6 == Year
#Column 2 == Short_name
#Just need to input the number of the row (1-13)

x <- 8


readxl::read_excel(here::here(  'data', 
                                paste(paste(FCT_QA[x,6],
                                            FCT_QA[x,2], sep = '_'), 'xlsx', sep = '.')), sheet = 1) %>%
  head()


#Customized saving FCT

NGAFCT <- readxl::read_excel(here::here(  'data', 
                                          paste(paste(FCT_QA[x,6],
                                                      FCT_QA[x,2], sep = '_'), 'xlsx', sep = '.')), 
                             sheet = 1) %>% mutate(FCT = 'NGAFCT')  %>% glimpse()

NGAFCT <- NGAFCT %>% rename(
  ref = 'REFID',
  code = 'Code', 
  foodgroup = 'Category',
  fooditemNGA = 'LocalName',
  fooditem = 'EnglishName',
  scientificName = 'ScientificName',
  fooditemFR = 'FrenchNames',
  ENERC1 = 'ENERC_kcal',
  ENERC2 = 'ENERC_kJ')


#Removing units from variable names

names(NGAFCT) <- sub("_g|_mcg|_mg", "", names(NGAFCT))

NGAFCT <- NGAFCT %>% rename_at(vars(16:25), funs(toupper)) %>%
  rename(VITA_RAE = 'VIT_A_RAE',
         VITB6 = 'VIT_B6',
         FIBGT = 'FIB')


NGAFCT <- NGAFCT %>% mutate_at(vars(9:37), funs(as.numeric)) 


NGAFCT <- NGAFCT %>%
  mutate(SOP = reduce(select(., #It is measuring total CHO, which includes fibre
                             'WATER', 'PROTCNT' ,'FATCE', 'CHOCDF',  'ASH'), `+`))

```
####9) South Africa FCT

ONLY PDF

####10) Tanzania FCT

ONLY PDF

####11) Uganda FCT

There are some variables that we want to keep but we are not going to rename. The fastest approach is to proceed as with KENFCT and cut into two part. Then, to modify separately both and then bind them. 

UGAFCT1 

To include 'ref' as the reference of where the food item is coming from, there are three variables that offer information of the reference source: named fct_source_*. We chose 'fct_souce_descr' because it the one that provided more information from the other three (however, all three remind available in the fct). 

UGAFCT2

2) Dry Matter (g): DM

5) lipid_total_g: FAT - It is stated that it was calculated by by gravimetric methods

7) fibre_g: FIBTG - No information is provided on how fibre was calculated. We are assuming that is total dietary fibre. 
TO BE REVIEWED

21 vit_a_iu: VITA_IU - Vitamin A in IU. This form of expressing VITA is not recommended. In the user-guides is acknowledged and it is adviced not to used, unless no other data is available. They also provided a conversion factor to VIT_RAE for plant and animal food sources. Since, retinol, beta- and alpha- carotene and cryptoxanthin are reported, VITA could be calculated following the proposed equation from FAO/INFOODS.

24) alpha_carot_mcg: CARTA (alpha-carotene, Unit: mcg). 

26) beta_crypt_mcg: CRYPXB (beta-cryptoxanthin, Unti: mcg).

NOTE: Ash is not been reported

```{r UGAFCT tagname }

#Uganda FCT - tagname standardization

#View FCT structure
#Creating a variable for the names of the data-set 
# it is composed by 2 columns on the FCT_QA dataset
#Column 6 == Year
#Column 2 == Short_name
#Just need to input the number of the row (1-13)

x <- 11


readxl::read_excel(here::here(  'data', 
                                paste(paste(FCT_QA[x,6],
                                            FCT_QA[x,2], sep = '_'), 'xlsx', sep = '.')), sheet = 1) %>%
  head()


#Customized saving FCT (1)

UGAFCT1 <- readxl::read_excel(here::here(  'data', 
                                          paste(paste(FCT_QA[x,6],
                                                      FCT_QA[x,2], sep = '_'), 'xlsx', sep = '.')), 
                             sheet = 1) %>% mutate(FCT = 'UGAFCT')  %>% select(1:13) %>% glimpse()

#Renaming some variables

UGAFCT1 <- UGAFCT1 %>% rename(
                      code = 'food_code', 
                      fooditem = 'food_description',
                      ref = 'fct_source_descr',
                      foodgroup = 'food_group')

#Customized saving FCT (2)

UGAFCT2 <- readxl::read_excel(here::here(  'data', 
                                          paste(paste(FCT_QA[x,6],
                                                      FCT_QA[x,2], sep = '_'), 'xlsx', sep = '.')), 
                             sheet = 1)   %>% select(14:39) %>% mutate(FCT = 'UGAFCT') %>% glimpse()

#Variable names acc. tagname

FCT11_tag <- c( 'WATER', 'DM', 'ENERC1', 
              'PROTCNT', 'FAT',  'CHOAVLDF', 'FIBTG',  
              'CA', 'FE',  'ZN', 'VITC', 'THIA', 'RIBF',
              'NIA', 'VITB6', 'FOL', 'FOLAC', 'FOLDF', 'FOLDFE', 
              'VITB12', 'VITA_IU', 'VITA_RAE', 'RETOL', 'CARTA', 'CARTB' , 'CRYPXB', 
              'FCT')

UGAFCT2 <- UGAFCT2 %>% rename_all( ~ FCT11_tag) 


UGAFCT <- bind_cols(UGAFCT1, UGAFCT2)

UGAFCT <- UGAFCT %>%
  mutate(SOP = reduce(select(.,
                             'WATER', 'PROTCNT' ,'FAT', 'CHOAVLDF','FIBTG'), `+`))






summary(UGAFCT$SOP)

```

####12) uFish FCT

This FCT needs only minor renaming since it has all components named using tagnames form FAO/INFOODS. Scientific name are available in the user-guide.

The table reported the ASFIS List of Species for Fishery Statistics Purposes, which refers to column 2 and 3. 

2) ISSCAAP: International Standard Statistical Classification of Aquatic Animals and Plants. Remind unchanged.

3) 3-Alpha: three_alpha - International code for fish.

5) Habitat: referring to fish habitat coded as F - Farmed and W - Wild. Remind unchanged.

6) State of food: food_state - referring to r - raw, c - cooked, p - processed/preserved. We used same code as UGAFCT while keeping it as close as the original uFish. 

8) EDIBLE...8: EDIBLE1 - Following the same rationale as in WAFCT

9) EDIBLE...9: EDIBLE2 - Following the same rationale as in WAFCT

For FAO/INFOODS tagname, I will change all at once for NGAFCT, uFish and uPulses.

```{r uFish tagname }

#uFish FCT - tagname standardization

#View FCT structure
#Creating a variable for the names of the data-set 
# it is composed by 2 columns on the FCT_QA dataset
#Column 6 == Year
#Column 2 == Short_name
#Just need to input the number of the row (1-13)

x <- 12


readxl::read_excel(here::here(  'data', 
                                paste(paste(FCT_QA[x,6],
                                            FCT_QA[x,2], sep = '_'), 'xlsx', sep = '.')), sheet = 4) %>%
  head()

#Customized saving FCT

uFish <- readxl::read_excel(here::here(  'data', 
                                           paste(paste(FCT_QA[x,6],
                                                       FCT_QA[x,2], sep = '_'), 'xlsx', sep = '.')), 
                              sheet = 4) %>% slice(2:516) %>% mutate(FCT = 'uFish')   %>% glimpse()

uFish <- uFish %>% rename(
  code = 'Food Item ID', 
  three_alpha = '3-Alpha',
  fooditem = 'Food name in English',
  food_state = 'State of food',
  ref = 'RefID',
  EDIBLE1 = 'EDIBLE...8',
  EDIBLE2 = 'EDIBLE...9',
  ENERC1 = 'ENERC(kcal)',
  ENERC2 = 'ENERC(kJ)')

names(uFish) <- sub("\\(.*?\\)", "", names(uFish))


uFish <- uFish %>% mutate_at(vars(8:168), funs(as.numeric)) 


uFish <- uFish %>%
  mutate(SOP = reduce(select(.,
                             'WATER', 'PROTCNT' ,'FAT', 'CHOAVLDF','FIBTG',  'ASH'), `+`))

summary(uFish$SOP)

```

####13) uPulses FCT

The structure is very similar to uFish and we will proceed in the same fashion. 

2) 'Country, region': origin - reference of the origin of the food item

4) Processing: food_state coded as r: raw, d: dried, p: processed.

5) Species/Subspecies: scientificName - In the User-guide are refered as the scientific names. Hence, to keep consistency,we have used the same variable names as in WAFCT for the scientific name.

6) Cultivar/Variety/Accession Name: variety - for simplicity we have kept the most common name of the three possible.

FATCE: Fat values derived from Soxhlet method (FATCE)

```{r uPulses tagname }

#uPulses FCT - tagname standardization

#View FCT structure
#Creating a variable for the names of the data-set 
# it is composed by 2 columns on the FCT_QA dataset
#Column 6 == Year
#Column 2 == Short_name
#Just need to input the number of the row (1-13)

x <- 13


readxl::read_excel(here::here(  'data', 
                                paste(paste(FCT_QA[x,6],
                                            FCT_QA[x,2], sep = '_'), 'xlsx', sep = '.')), sheet = 4) %>%
  head()

#Customized saving FCT

uPulses <- readxl::read_excel(here::here(  'data', 
                                         paste(paste(FCT_QA[x,6],
                                                     FCT_QA[x,2], sep = '_'), 'xlsx', sep = '.')), 
                            sheet = 4) %>%  slice(2:177) %>% mutate(FCT = 'uPulses')   %>% glimpse()

uPulses <- uPulses %>% rename(
  code = 'FoodID', 
  origin = 'Country, region',
  fooditem = 'Food name in English',
  food_state = 'Processing',
  scientificName = 'Species/Subspecies',
  variety = 'Cultivar/Variety/Accession Name',
  ref = 'BiblioID',
  ENERC1 = 'ENERC(kcal)',
  ENERC2 = 'ENERC(kJ)')


names(uPulses) <- sub("\\(.*?\\)", "", names(uPulses))

uPulses <- uPulses %>% mutate_at(vars(8:68), funs(as.numeric)) 


uPulses <- uPulses  %>%
  mutate(SOP = reduce(select(.,
                             'WATER', 'PROTCNT' ,'FATCE', 'CHOAVLDF','FIBTG',  'ASH'), `+`))


summary(uPulses$SOP)

```

##Conclusion

1) Energy: Most of the FCT used the ENERC values and conversion factor as established by FAO/INFOODS. Most of the differences were including calculation for kcal and/or kJ. Including ALC or not into the calculation. Mozambique kcal were calculated as kJ and applying a conversion factor (Energy, kJ/4,184). Gambia did not account for fibre and the conversion factor used for CHO was 3.75 (due to the calculation is based on monosaccharide equivalents). In addition, Tanzania used total carbohydrates, hence fibre was included into the CHO fraction (not accounting for differences in fibre conversion factors). Three FCT (eth, nga, uga) did not report the method used to calculate the energy content. This variable will be (re-)calculated following the equations:

Eq.1: ENERC - (kcal/100 g EP) = PROTCNT (g/100 g EP) x 4 + FAT (g/100 g EP) x 9 + CHOAVLDF (g/100 g EP) x 4 + FIBTG (g/100 g EP) x 2 + ALC (g/100 g EP) x 7]

Eq.2: ENERC - (kJ/10- g EP) = PROTCNT (g/100 g EP) x 17 + FAT (g/100 g EP) x 37 + CHOAVLDF (g/100 g EP) x 16 + FIBTG (g/100 g EP) x 8 + ALC (g/100 g EP) x 29]

2) Protein: All FCT that reported used similar approach. It was calculated using total nitrogen (NT) and either a general conversion factor (6.75) or specific conversion factors (XN). If NT information is available for those that used general conversion factor, we could re-calculate using the conversion factors. 

Eq.3: PROTCNT - (g/100 g EP) = nitrogen conversion factor (XN) × total nitrogen (NT) (g/100 g EP)

3) Fat: FAT - is the most common used and it refers to fat calculated using mixed solvent extraction. This was not initially compiled in FCT_QA.  

4) Carbohydrates: Most of the FCT calculated CHO as CHOAVLF (carbohydrates available) as per FAO/INFOODS recommendation with slight variability of the presence or absence of ALC (alcohol). Tanzania reported total carbohydrates (CHOTDF). While, Gambia FCT accounted for available CHO by summation and expressed as monosaccharide equivalents (CHOAVLM). MAFOODS reported total CHO by summation (CHOCSM/CHOTSM?). ETHFCT reported CHOT. According to FAO/INFOOD the preferred reporting is CHOAVL followed by CHOAVLF.  

5) Fibre: Most of the dataset reported total fibre (FIBTG), while MAFOODS reported FIBC (and marked as italic when FIBTG was available) and GMBFCT that reported FIBTS (the sum of non-digestible polysaccharides and lignin). According to the FAO/INFOODS the preferred reporting method is FIBGT followed by FIBTS. While FIBC should be phased out. 

