---
title: "Fortification_checks"
author: "LuciaSegovia"
date: "8/17/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Fortification

Fortification calculations:

Maize grain:

60mg FE /kg DW
30mg ZN/ Kg DW

Source: spreadsheet-unfortified-crops

10g WATER/ 100g EP

Source: USDA

6mg/ 100g DW * (100-10g WATER/ 100g ) / 100 = 

FE in maize grain 6* (100-10)/ 100=5.4 mg FE / 100g EP

ZN in maize grain 3* (100-10)/ 100=2.7 mg ZN / 100g EP



Source: FAO/INFOODS unit conversion



http://ffinetwork.org/country_profiles/country.php?record=130



```{r cars}
summary(cars)
```

## Including Plots

I checked vitamin A values for self-declared fortified maize products (‘fortified in Vitamin A’) 

VITA (RAE) mean = 395mcg 
VITA (RAE) min. = 117mcg 

Then, for self-declared unfortified maize (‘unfortified’)

VITA (RAE) mean = 0.812mcg 
VITA (RAE) max. = 13mcg 


USDA value for corn 

1) Corn grain, white: VITA (RAE) = 0mcg

source: https://fdc.nal.usda.gov/fdc-app.html#/food-details/168920/nutrients


2) Corn grain, yellow:  VITA (RAE) = 11mcg 

source: https://fdc.nal.usda.gov/fdc-app.html#/food-details/170288/nutrients

We have confirmed the values for corn flour, unfortified as well. 


```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
