---
title: "Fortification_checks"
author: "LuciaSegovia"
date: "8/17/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Fortification

http://ffinetwork.org/country_profiles/country.php?record=130



```{r cars}
summary(cars)
```

## Including Plots

I checked vitamin A values for self-declared fortified maize products (‘fortified in Vitamin A’) 

VITA (RAE) mean = 395mcg 
VITA (RAE) min. = 117mcg 

Then, for self-declared unfortified maize (‘unfortified’)

VITA (RAE) mean = 0.812mcg 
VITA (RAE) max. = 13mcg 


USDA value for corn 

1) Corn grain, white: VITA (RAE) = 0mcg

source: https://fdc.nal.usda.gov/fdc-app.html#/food-details/168920/nutrients


2) Corn grain, yellow:  VITA (RAE) = 11mcg 

source: https://fdc.nal.usda.gov/fdc-app.html#/food-details/170288/nutrients

We have confirmed the values for corn flour, unfortified as well. 


```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
