---
title: "FCT_CleaningSOP"
author: "LuciaSegovia"
date: "8/4/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

We started with West Africa FCT, the data set is provided in excel with XX sheet. Corresponding to metadata and variables description, FCT (short and long version), FCT statistics, recipe calculation, etc. This FCT follows the FAO guidelines for FCT compilation. 

We need to import the FCT food item list sheet (long version) which is located in the sheet 5. Then, some cleaning and renaming of variables was needed. 

DISCUSSION POINT: Variables names. Are we going to follow FAO standards??

Yes, we will use INFOODS tagnames as standard names

To make the code consistent for all the datasets. I will recode the name of the food items in English as 'fooditem'. If there is another language available, the code will be 'fooditem' and the language code (upper-case) i.e. 'fooditemFR'. This is important because if we use 'name' is a very generic variable name, and it could cause confusion and bugs in our code. 


```{r cars}
summary(cars)
```


The next step would be to gather the table back, to separate units from the components and to match the components. Particularly for those of interest for MAPS. 

Additionally, some other general information can be collected from the datasets themselves, such as food categories, number of food items, components included and so forth. 

This part is carried out after the section b, because it is going to be done in parallel with fct cleaning

For extracting component list (names), we need to create a list. Then, we can count (lenght) the amount of food categories by unlisting. 

NOTE: I am not replacing the food categoies for WAFCT because when I pulled manually I only extracted the English names, and I attached the food group number (code).

  

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
