---
title: "2019_WAFCT_Cleaning"
author: "LuciaSegovia"
date: "8/11/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)

```

###Exploring data in the xlsx file

   
```{r exploring sheets}



```

We can see that there are 12 sheets, we are going to use several of the sheets, such as 5, 7, 8, 10.

We are going to start with the sheet 5, in this table, components are accompanied by their tagnames. 

We are going to standardize variables names as descibed above for food items names and code and for food components tagnames (FAO/INFOODS). Additionally, food groups will be standardized using FoodEx2 classification system.

UPDATE: Add rename to Energy variables folowing the standard that it is set below (MAFOODS) 
 +ENERC1 = ENERC...10 = Energy in kcal; 
 +ENERC2 = ENERC...10 = Energy in kJ).
 
  -FOL: some of the values are reported as FOLSUM. This tagname is recommended to be discontinued. 

  -PHYYCPP: some values were reported as PHYTCPPD or PHYCPPI.

```{r WAFCT standard variables}


```

NOTE: The problem with converting to numeric is that number between [] are removed (as NA). 
