---
title: "IHS Food <atching"
author: "Segovia De La Revilla, Lucia"
date: "20/09/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Updated: 2023-05-31

# Introduction

Contains information on the food matching for Malawi household surveys, including IHS4 and IHS5.

# IHS 4
**Change log**
V.2.0 - updated on  2023-05-31: Fixed items with not matches on any of the FCTs used in MAPS tool. 

#Important

Thobwa: 24230.03.03 - maize- and millet-fermented beverage, change description in dictionary & found a match for millet-fermented beverage on WA19.
  
Maize ufa madeya (bran flour): 39120.04.01 - maize bran, flour, raw --> No information, changing to whole grain flour (23120.03.01)
  
Gathered wild green leaves: 1214.06 - blackjack leaves, raw - recoded (1290.9.16)
  
Squash (Sobo drink concentrate):  24490.01 - sugar sweetned beverage, concentrate (New code)

Sun Dried fish (Small Variety):  1505.03 -  barb, silver, dried, raw (New code)
  
Fresh fish (Small Variety):  1501.06 -  barb, silver, fresh, raw - recode (1501.11)
  
Groundnut fresh (unshelled): 142.03 - peanuts, fresh, unshelled, raw. No information on fresh peanuts, hence we need to adjust the quantity from fresh consumption to dry E.g., 100g of fresh weight will be 100*(100-50)/(100-10), and matched to dry peanuts. This is important to not over-estimate the nutrient content from the food consumption.
  
Freezes (flavoured ice):  23670.01.02 - flavoured ice - recoded ("22270.06")
  
Smoked fish (Small Variety):  1505.06 -  barb, silver, smoked, raw (New code)
  
Maheu: 24230.03.02 - maize-fermented beverage - recoded ("24310.02.01")
  
Chambiko - soured milk: 22290.01 - sour milk, traditional - recoded("22290.05")
  
Sun Dried fish (Large Variety): 1505.02 - tilapia, dried, raw (New code)
  
Smoked fish (Large Variety):  1505.05 - tilapia, smoked, raw (New code)
  
Smoked fish (Medium Variety):  1505.05 -  tilapia, smoked, raw (New code)

Hot sauce (Nali, etc.): 1699.06 - tabasco, sauce - recode (F1232.04)
  
Termites, other insects (eg Ngumbi, caterpillar): 21170.92.02 - edible insect adults, fresh, raw - recode (21183.02) (termite dry as in MW19).
  
Cassava - roasted (vendor): 1520.01.03 - cassava root, roasted - recode (1520.01.05) Roasted was not available
  
Rosted sweet potatoes: 1530.05 - sweet potato, white-flesh, roasted (New code) - To be reviewed.
  
  # Secondary
  23140.03.02
  Breakfast cereal
  breakfast cereals, muesli mix
  
  
  1214.05
  Other cultivated green leafy vegetables
  jews mallow leaves, raw
  	
Other cultivated green leafy vegetables: 1215.03 roselle leaves, raw - recode()


Citrus ÃƒÂ¯Ã‚Â¿Ã‚Â½ naartje, orange, etc.: 1324.02 - mandarins, raw (new code)

Maize - boiled or roasted (vendor): 1290.01.02 - maize, green, roasted (new code)

Sun Dried fish (Large Variety): 1505.01 - catfish, african, dried, raw. Catfish from fresh to dried. --> Important due to the water content. Adjust the quantity to fresh E.g., 100g of drt weight will be qty*(100-DW)/(100-FW) Water values based on catfish WA19(09_060) - 78g (FW) and MW19(MW03_0048) - 21g (DW)

Other poultry - guinea fowl, doves, etc.: 21170.01.01 - pigeon meat, fresh, raw (new code)


F0623.03
Zikondamoyo/Nkate
banana, fritters

1505.04
Smoked fish (Large Variety)
catfish, african, smoked, raw

1505.04
Smoked fish (Medium Variety)
catfish, african, smoked, raw

F1061.02
Meal eaten at restaurant (vendor)
chicken meat, fresh, deep-fried

21170.92.03
Termites, other insects (eg Ngumbi, caterpillar)
edible insect larvae, fresh, raw

1501.04
Fish (vendor)
catfish, african, fresh, grilled


21119.01.01
Small animal ÃƒÂ¯Ã‚Â¿Ã‚Â½ rabbit, mice, etc.
rat meat, fresh, raw


F1061.02
Meat (vendor)
chicken meat, fresh, deep-fried

1530.07
Rosted sweet potatoes
sweet potatoes, orange-flesh, roasted

F0623.02
Jam, jelly
jelly, fruits

Boiled sweet potatoes:  1530.06 - sweet potatoes, orange-flesh, boiled

## IHS5-foodid matching to Genus

There are some items that should be removed:

 1) [106] There are two codes for identical item = rice. We amended genus to account for imported and local rice production hence, differentiate between the two. Unless, it is specified, we will sort for local when possible. 
 2) [510] It is a redundant code, as mice is actually rabbit meat. We should remove and use only rabbit until we found mice. 
 3) [816] Remove sweets, chocolate with milk because acc. to Gareth inputs that's hardly found in Malawi. We identified an issue that with the other code that should be solved at FCT level 
 4) [832] Remove orange-sweet potato. It also has an issue that it should be fixed at FCT level.
 5) [836] A typo in food genus that was solved in version 2.5 samosa and banana cake sharing the same id, was the cause of this error. Now it's solved. Hence it doesn't need to be removed any more. 
 
IHS5 - genus matches were sent to Gareth Osman to double-check that the assumptions made were consistent with Malawi reality. 

Below are the comments to be solved:

 1) 118 - 'MAIZE UFA RAW MADEYA (bran flour - unprocessed), code 118, is missing. --> Fixed
 2) 412 - Tinned vegetables (specify) change to Tinned beans. = Tinned, beans (MW02_0003) --> Fixed
 3) 106 - rice was duplicated --> Fixed. 
 4) 829 - restaurant meals ---> Nsima (MW01_0031 - F1232.05) - check min values for this recipe, chicken (MW03_0010 - F1061.01), rape leaves, boiled (MW04_0020*Retention factor), change to amaranth leaves, boiled (MW04_0010 - ) --> Fixed but w/o retention factors
 5) 816 - Change compo to candies
 6) 708 - Add infant formula data
 
-Recalculate min values for all sweet potato using Edward:

-Provisional fix to create OFSP version

 6) 832 and 831 - Add b) version w/ orange: copy and paste values and 
 7) 204 - OFSP
 8) 203 - WFSP
 
Changed value of VITA_RAE of (831) white boiled sweet potato to match raw white sweet potato adjusted w/ retention factor. We did not do a water adjustement because the water content difference in water was <10%, hence according to FAO/ INFOODS Guidelines there is no need to water adjust 
[2.213000*(100-76.00/100-77.28)]
 
Make a note to fix all recipes in MAFOODS.

- Check ref for small fish:

 9)  5023        Sun Dried fish (Small Variety)     
 10) 5033        Fresh fish (Small Variety)  
 11) 5123        Smoked fish (Small Variety) 

- Check infant feeding formula


```{r ihs5-quality-check}

```

## FCT-IHS5 matching to Genus


Keep 

* 1214.05 = NA
* 1214.06 = NA
- 1215.02 = MW04_0011
- 1215.03 = MW04_0021
- 1219.01.01 = MW04_0014
- 1322.01 = MW05_0014
- 1323.01 = MW05_0018
- 1359.9.01 = MW05_0013
- 1359.9.02 = MW05_0005
- 1501.02 = 91010          
- 1501.05 = 8010           
- 1505.01 = MW03_0020
- 1505.02 = MW03_0031
- 1505.04 = MW03_0023
- 1505.05 = MW03_0047
* 1699.04 = NA
- 1699.05 = 13002
* 21111.01.02 = NA
* 21116.02 = NA
- 21121.04 = 15073
- 21170.92.02 = MW03_0069  (??)
- 21170.92.03 = MW03_0009
- 21397.01.01 = 4025
- 21431.01 = MW05_0012
- 2161.01 = 11_009
- 2162.01 = 11_003
- 21631.01.01 = 9013
- 22251.01.01 = 6005
- 23140.03.01 = 1009
* 23140.03.02 = NA
- 23991.01.01 == All??
- 24310.02.01 = 12_002
- 39120.04.01 = MW01_0035
- F0020.01 = MW01_0004
- F0020.02 = MW01_0003
* F0022.06 = NA
- F0623.03 = MW01_0061
- F0666.01 == REMOVE ??



food_genus_id - ref_fctcode that should be removed:

01520.01.03 - MW01_0010 # roasted vs boiled cassava
1235.01 - MW04_0034 # raw vs boiled pumpkin ??
1290.01.01 - MW01_0040 #raw vs boiled green maize ?? 
1290.01.02 - MW01_0040 #roasted vs boiled green maize ??
142.03 - MW02_0015 #fresh vs boiled peanuts ??
142.04 - 8006 # wrong name 
1501.04 - MW03_0030 # tilapia vs catfish
1530.07 - 5010 #white vs orange flesh sweet potato
1532.01 - 07_063 #sardines tinned vs meat tinned
1699.06 - 13031 #tabasco vs tomato ketchup *has different genus
21119.01.01 - MW03_0067 #rat/mice vs rabbit
21121.02 - MW03_0011 # chicken meat vs minced chicken meat
21170.01.01 - MW03_0065 #pigeon vs quail ??
21170.01.02 - MW03_0065 #guinea fowl vs quail ??
23110.01 - 01_043 #unrefined vs refined wheat flour
23161.01.01 - MW01_0058 #imported vs local rice
23670.01.01 - 13_021 #chocolate vs candies
23670.01.02 - 12_012 #flavoured ice - carbonated drink  ??
2413.01 - 22007 #gin vs wine
24230.03.03 - 12_002 #sorghum and maize beer vs maize beer ??
24310.04.01 - 140001 #sorghum beer vs beer
F0022.01 - MW01_0041 #buns vs scones
F0623.02 - 13_023 #jelly vs jam


24310.02.01 - 140001 #beer
F0022.06  -  15026 #vege samosa
23670.01.01 - 13_021 sweets, chocolate, milk

Items that food item in the fct_ihs5 doesn't match to the genus code

- 816 - 23670.01.01 - WAFCT - 13_021 sweets, chocolate, milk --> should be candies, hard
- Roasted sweet potatoes: 1530.07 - AHHA - *sweet potato, white, eaten without skin, grilled --> should be orange flesh roasted/grilled sweet potato.

Items that doesn't match and/or are redundant
- 21119.01.01 - MAFOODS - MW03_0067 - rabbit meat --> should be mice, otherwise remove.
- 23161.01.01 - Rice, imported. 
- F1232.08 - Baking soda --> Substitution of baking powder (and yeast)

#### List of non-available items in WA19 compared w/ mwi_genus

F0022.04 = scones 
118.02 - only pearl millet
23120.03.01 = maize, flour, white, unrefined, non-fermented, raw
F0022.02 = dough, sweet, fried?
Check type of plantain
Check type of sorghum
1701.03
141.02
Check peanuts - all shelled
142.02 - check shelled?
142.05 - check shelled?
21111.01.01 - check it said w/ bones
21170.01.03
1212.02
Check leaves - fresh or dried??
1214.03
1270.01
(1341.01) - apple here is w/ skin
(1316.01) - check mangoes here conservative pale flesh
21700.02.01 - margarine only fortified
1802.01
2899.01.01
