
# Nutrient Conversion Table for Malawi Integrated Household Survey


The scripts described here can be used for generating a Nutrient Conversion Table (NCT) to be use with the Malawi Integrated Household Survey (IHS) round 4 (2016-2017), and round 5 (2019-2020) with slight modifications. 

This protocol serves as documentation of the food matches that are performed in the back-end of the MAPS tool{LINK}. 

Last Updated: 2024-08-12


**Change log**
V.2.0.0 - updated on  2023-05-31: Fixed items without matches on any of the FCTs used in MAPS tool, changed to better matches in the FCTs.


# Introduction

The document contains information on the food matching for Malawi household surveys, including IHS4 () and IHS5 ().


# Updating food matches to be performed within the MAPS tool.

 - Checking excluded in IHS5 script (ihs5_genus.R)
 - Here's the changes that were made however no serious issue or good reason seemed recoded (beside data needs)
 - Not changing them back.


```{r}

# Checking excluded in IHS5 script (ihs5_genus.R)
food_list %>% 
filter(ID_3 %in% # Citrus and orange sweet potato from boiled version
        c("1530.06", "1321.02", "1322.02", "1324.02", 
          # Changed green and roasted maize to boiled 
          "1290.01.01", "1290.01.02" , "1290.01.03",
          # Yeast and baking powders 
              "F1232.07", "F1232.06",
        ))  %>% View()

```

NOTE: Ca values in baking powder ("F1232.07") are high, specially **extremely** although plausible in US19(18369)


```{r}

# Checking yeast and baking powder compo values
fct_dict %>% 
  filter(ID_3 %in%  
           c( "F1232.07", "F1232.06"))  %>%
  select(source_fct, ID_3, fdc_id,  food_desc,  CAmg, SEmcg, VITA_RAEmcg, VITB12mcg)
  View()

```

Some food needed a better match, for instance, beef was matched to "with bones" however, the edible portion used was 1. Hence, we change it to a one-to-many match using beef with three fat content and w/o bones. 

beef, lean, with bones, fresh, raw (21111.01.01) --> Added beef w/o bones, including ground raw. c("21111.02.03", "21184.01.02", "21111.02.01", "21111.02.02")


# Food items with better matches


```{r}

# Food items with better matches
# groundnuts (to dry and shelled) - Corrections should be done at Edible Portion
food_list$ID_3[food_list$code %in% c("311", "313")] <-  "142.01" 
food_list$ID_3[food_list$code %in% c("505")] <-  "21116.03" 
food_list$ID_3[food_list$ID_3 %in% c("21170.01.01")] <-  "21170.01.03" 
food_list$ID_3[food_list$ID_3 %in% c("1501.11")] <-  "1501.10" 
food_list$ID_3[food_list$ID_3 %in% c("1501.05")] <-  "1503.03" 
food_list$ID_3[food_list$ID_3 %in% c("21121.02")] <-  "21121.03" 
#Fixing samosa vs banana cake id typo issue (see MAPS_Dictionary-Protocol update v2.5)
food_list$ID_3[food_list$ID_3 == "F0022.06" & food_list$code == "836" ] <- "F0022.07"
#Meal eaten at restaurant (vendor) c(21116.02, 21121.04) -->  F1061.01 
#TODO: Generate a recipe c(F1061.01,F1232.05 )
food_list$ID_3[food_list$code == "829"] <- "F1061.01,F1232.05" 


```


106 - Rice: We changed from multiple (imported & local, white) to only rice grain, local, white, dried, raw (23161.02.01), because it is the only reported in FAO SUA (2015-2017).

# Food items reported in the IHS4 without any matches

Thobwa: 24230.03.03 - maize- and millet-fermented beverage, change description in dictionary & found a match for millet-fermented beverage on WA19.
  
Maize ufa madeya (bran flour): 39120.04.01 - maize bran, flour, raw --> No information, changing to whole grain flour (23120.03.01)
  
Gathered wild green leaves: 1214.06 - blackjack leaves, raw - recoded (1290.9.16)
  
Squash (Sobo drink concentrate):  24490.01 - sugar sweetned beverage, concentrate (New code)

Sun Dried fish (Small Variety):  1505.03 -  barb, silver, dried, raw (New code)
  
Fresh fish (Small Variety):  1501.06 -  barb, silver, fresh, raw - recode (1501.11)
  
[313] - Groundnut fresh (unshelled): 142.03 - peanuts, fresh, unshelled, raw. No information on fresh peanuts, hence we need to adjust the quantity from fresh consumption to dry E.g., 100g of fresh weight will be 100*(100-50)/(100-10), and matched to dry peanuts. This is important to not over-estimate the nutrient content from the food consumption.

```{r}
# Checking the values of groundnuts in MW19

fct_dict %>% filter(fdc_id %in% c("MW02_0015", "MW02_0014")) %>% 
                      dplyr::select(fdc_id, food_desc, WATERg, SEmcg)


```

  
Freezes (flavoured ice):  23670.01.02 - flavoured ice - recoded ("22270.06")
  
Smoked fish (Small Variety):  1505.06 -  barb, silver, smoked, raw (New code)
  
Maheu: 24230.03.02 - maize-fermented beverage - recoded ("24310.02.01")
  
Chambiko - soured milk: 22290.01 - sour milk, traditional - recoded ("22290.05")
  
Sun Dried fish (Large Variety): 1505.02 - tilapia, dried, raw (New code)
  
Smoked fish (Large Variety):  1505.05 - tilapia, smoked, raw (New code)
  
Smoked fish (Medium Variety):  1505.05 -  tilapia, smoked, raw (New code)

Hot sauce (Nali, etc.): 1699.06 - tabasco, sauce - recode (F1232.04)
  
Termites, other insects (eg Ngumbi, caterpillar): 21170.92.02 - edible insect adults, fresh, raw - recode (21183.02) (termite dry as in MW19).
  
Cassava - roasted (vendor): 1520.01.03 - cassava root, roasted - recode (1520.01.05) Roasted was not available
  
Rosted sweet potatoes: 1530.05 - sweet potato, white-flesh, roasted (New code) - To be reviewed.
  
## Food items reported in the IHS4 with incomplete matches in multiple matches

Breakfast cereal (23140.03.02) - breakfast cereals, muesli mix - recode (23140.03.01)
  
Other cultivated green leafy vegetables (1214.05) - jews mallow leaves, raw - recode (1290.9.15)
  	
Other cultivated green leafy vegetables (1215.03) roselle leaves, raw - recode(1290.9.10)

Citrus ÃƒÂ¯Ã‚Â¿Ã‚Â½ naartje, orange, etc.: (1324.02) mandarins, raw (new code)

Maize - boiled or roasted (vendor): 1290.01.02 - maize, green, roasted (new code)

Sun Dried fish (Large Variety): 1505.01 - catfish, African, dried, raw. Catfish from fresh to dried. --> Important due to the water content. Adjust the quantity to fresh E.g., 100g of dry weight will be qty*(100-DW)/(100-FW) Water values based on catfish WA19(09_060) - 78g (FW) and MW19(MW03_0048) - 21g (DW)

Other poultry - guinea fowl, doves, etc.: 21170.01.01 - pigeon meat, fresh, raw (new code)

Zikondamoyo/Nkate: (F0623.03) banana, fritters - New code

Smoked fish (Large Variety): (1505.04) catfish, african, smoked, raw (new code)

Smoked fish (Medium Variety): (1505.04) catfish, african, smoked, raw (new code)

Meal eaten at restaurant (vendor): (F1061.02) chicken meat, fresh, deep-fried (new code)

Termites, other insects (eg Ngumbi, caterpillar): (21170.92.03) edible insect larvae, fresh, raw - recode (21183.03) & new code

Fish (vendor): (1501.04) catfish, african, fresh, grilled - recoded (1507.12)

Small animal ÃƒÂ¯Ã‚Â¿Ã‚Â½ rabbit, mice, etc. (21119.01.01) rat meat, fresh, raw (New code)

Meat (vendor): (F1061.02) chicken meat, fresh, deep-fried (new code)

Jam, jelly: (F0623.02) jelly, fruits - recode (23670.01.05) & New code

#### Orange fleshed sweet potato

We are using an allocation factor of .162 (16.2%), according to G.O. analysis of the coverage of OFSP in Malawi (Osman et al. *unpublished*).

# Roasted sweet potatoes: (1530.07) sweet potatoes, orange-flesh, roasted
# Boiled sweet potatoes:  (1530.06) sweet potatoes, orange-flesh, boiled

## Checking Vitamin A values (VITA_RAEmg)

- Items without values: 	

cabbages, head, fresh, raw (1212.01) - New code: could be improved with an average of cabbages.

blackjack leaves, raw (1290.9.16) -> Adding more items to the list: following IHS5:
 1. leaves, cat's whiskers, raw, cleome gynadra  (1290.9.07)
 2. jew's mallow leaves, raw (1290.9.15)
 

sugar sweetned beverage, concentrate (24490.01) - Assumed zero (no need to change because multple match)

# barb, silver, smoked, raw (1505.06)

north african catfish, fresh, raw (1501.02) - recode upsidedown catfish (1501.09) bc it is the same specie as in the MAFOODS. 
tilapia, fresh, raw (1501.05) - recode to fillet ("1503.08", "1503.03")

maize, green, roasted (1290.01.02) - Not completed bc multi-matched

cassava root, roasted (1520.01.05) - Not completed bc multi-matched


## Checking Iron values (FEmg) & Zinc (ZNmg)

Small animal ÃƒÂ¯Ã‚Â¿Ã‚Â½ rabbit, mice, etc.: (21119.01.01) rat meat, fresh, raw - Not completed bc multi-matched

#EDIBLE: SUGAR CANE
Sugar Cane: (1802.01) sugar canes - recode (1802.02) cane sugar juice. We need to adjust the quantity consumed to account for the difference in the edible portion. 

## IHS5-foodid matching to Genus

There are some items that should be removed:

 1) [106] There are two codes for identical item = rice. We amended genus to account for imported and local rice production hence, differentiate between the two. Unless, it is specified, we will sort for local when possible. 
 2) [510] It is a redundant code, as mice is actually rabbit meat. We should remove and use only rabbit until we found mice. 
 3) [816] Remove sweets, chocolate with milk because acc. to Gareth inputs that's hardly found in Malawi. We identified an issue that with the other code that should be solved at FCT level 
 4) [832] Remove orange-sweet potato. It also has an issue that it should be fixed at FCT level.
 5) [836] A typo in food genus that was solved in version 2.5 samosa and banana cake sharing the same id, was the cause of this error. Now it's solved. Hence it doesn't need to be removed any more. 
 
IHS5 - genus matches were sent to Gareth Osman to double-check that the assumptions made were consistent with Malawi reality. 

Below are the comments to be solved:

 1) 118 - 'MAIZE UFA RAW MADEYA (bran flour - unprocessed), code 118, is missing. --> Fixed
 2) 412 - Tinned vegetables (specify) change to Tinned beans. = Tinned, beans (MW02_0003) --> Fixed
 3) 106 - rice was duplicated --> Fixed. 
 4) 829 - restaurant meals ---> Nsima (MW01_0031 - F1232.05) - check min values for this recipe, chicken (MW03_0010 - F1061.01), rape leaves, boiled (MW04_0020*Retention factor), change to amaranth leaves, boiled (MW04_0010 - ) --> Fixed but w/o retention factors
 5) 816 - Change compo to candies
 6) 708 - Add infant formula data
 
-Recalculate min values for all sweet potato using Edward:

-Provisional fix to create OFSP version

 6) 832 and 831 - Add b) version w/ orange: copy and paste values and 
 7) 204 - OFSP
 8) 203 - WFSP
 
Changed value of VITA_RAE of (831) white boiled sweet potato to match raw white sweet potato adjusted w/ retention factor. We did not do a water adjustment because the water content difference in water was <10%, hence according to FAO/ INFOODS Guidelines there is no need to water adjust (`r round(2.213000*(100-76.00/100-77.28), 2)`)
 
Make a note to fix all recipes in MAFOODS.

- Check ref for small fish:

 9)  5023        Sun Dried fish (Small Variety)     
 10) 5033        Fresh fish (Small Variety)  
 11) 5123        Smoked fish (Small Variety) 

- Check infant feeding formula


```{r ihs5-quality-check}

```

## FCT-IHS5 matching to Genus


Keep 

* 1214.05 = NA
* 1214.06 = NA
- 1215.02 = MW04_0011
- 1215.03 = MW04_0021
- 1219.01.01 = MW04_0014
- 1322.01 = MW05_0014
- 1323.01 = MW05_0018
- 1359.9.01 = MW05_0013
- 1359.9.02 = MW05_0005
- 1501.02 = 91010          
- 1501.05 = 8010   --> Recoded        
- 1505.01 = MW03_0020
- 1505.02 = MW03_0031
- 1505.04 = MW03_0023
- 1505.05 = MW03_0047
* 1699.04 = NA
- 1699.05 = 13002 --> Excluded (extreme Ca values See KE18 docu)
* 21111.01.02 = NA
* 21116.02 = NA
- 21121.04 = 15073 --> recoded
- 21170.92.02 = MW03_0069  (??)
- 21170.92.03 = MW03_0009
- 21397.01.01 = 4025
- 21431.01 = MW05_0012
- 2161.01 = 11_009
- 2162.01 = 11_003
- 21631.01.01 = 9013
- 22251.01.01 = 6005 --> recoded
- 23140.03.01 = 1009
* 23140.03.02 = NA
- 23991.01.01 == All??
- 24310.02.01 = 12_002
- 39120.04.01 = MW01_0035
- F0020.01 = MW01_0004
- F0020.02 = MW01_0003
* F0022.06 = NA
- F0666.01 == REMOVE ??


## Fixing items


### Done - Reviewed and changed

1520.01.03 - MW01_0010 # roasted vs boiled cassava
1235.01 - MW04_0034 # raw vs boiled pumpkin 
1290.01.01 - MW01_0040 #raw vs boiled green maize
1290.01.02 - MW01_0040 #roasted vs boiled green maize 
142.03 - MW02_0015 #fresh vs boiled peanuts 
1501.04 - MW03_0030 # tilapia vs catfish
21119.01.01 - MW03_0067 #rat/mice vs rabbit
21121.02 - MW03_0011 # chicken meat vs minced chicken meat
21170.01.01 - MW03_0065 #pigeon vs quail 
21170.01.02 - MW03_0065 #guinea fowl vs quail
23161.01.01 - MW01_0058 #imported vs local rice 
F0022.01 - MW01_0041 #buns vs scones
F0623.03 = MW01_0061  #This item is Steamed maize & banana snack, (Mkate) while the proposed is a banana fritter which is not the same, different food item in MW19



#### To be checked:


142.04 - 8006 # wrong name 

1530.07 - 5010 #white vs orange flesh sweet potato
1532.01 - 07_063 #sardines tinned vs meat tinned
1699.06 - 13031 #tabasco vs tomato ketchup *has different genus
23110.01 - 01_043 #unrefined vs refined wheat flour

23670.01.01 - 13_021 #chocolate vs candies
23670.01.02 - 12_012 #flavoured ice - carbonated drink  ??
2413.01 - 22007 #gin vs wine
24230.03.03 - 12_002 #sorghum and maize beer vs maize beer ??
24310.04.01 - 140001 #sorghum beer vs beer

F0623.02 - 13_023 #jelly vs jam


24310.02.01 - 140001 #beer
F0022.06  -  15026 #vege samosa
23670.01.01 - 13_021 sweets, chocolate, milk

Items that food item in the fct_ihs5 doesn't match to the genus code

- 816 - 23670.01.01 - WAFCT - 13_021 sweets, chocolate, milk --> should be candies, hard
- Roasted sweet potatoes: 1530.07 - AHHA - *sweet potato, white, eaten without skin, grilled --> should be orange flesh roasted/grilled sweet potato.

Items that doesn't match and/or are redundant
- 21119.01.01 - MAFOODS - MW03_0067 - rabbit meat --> should be mice, otherwise remove.
 
- F1232.08 - Baking soda --> Substitution of baking powder (and yeast)

#### List of non-available items in WA19 compared w/ mwi_genus

F0022.04 = scones 
118.02 - only pearl millet
23120.03.01 = maize, flour, white, unrefined, non-fermented, raw
F0022.02 = dough, sweet, fried?
Check type of plantain
Check type of sorghum
1701.03
141.02
Check peanuts - all shelled
142.02 - check shelled?
142.05 - check shelled?
21111.01.01 - check it said w/ bones
21170.01.03
1212.02
Check leaves - fresh or dried??
1214.03
1270.01
(1341.01) - apple here is w/ skin
(1316.01) - check mangoes here conservative pale flesh
21700.02.01 - margarine only fortified
1802.01
2899.01.01
